### 2019-07-18 manuscript/response.tex

\begin{document}

\hfill \today

Dear Dr. Schoen and Reviewers,

\medskip
Thank you very much for your close reading of our manuscript and supportive comments. We appreciate your constructive suggestions, each of which we address below. You will find a great difference in the number of models that we have fitted for the revision based on reviewer suggestions (originally 12, now 29, with a special supplementary figure to enable easy visualization). There are two main reasons for this sizable increase.

@@

\begin{quote}
--beginning of the Discussion is perhaps unnecessary -- it's more like Introduction material
\end{quote}

The discussion was heavily edited. However, we consider that it is important to refresh some of the introductory key points due to the numerous and complex questions that are being addressed with SSE models. 
 % TODO %We can defend this on account of repeating some of the points because of numerous and complex  questions; aimed to prevent flipping pages back
% Fixed

\begin{quote}
--are the results from the analyses that include diploidization coherent?
I'm not sure how to phrase this -- I guess I'm asking if there's any suggestion that that parameterization is indeed capturing the biological phenomenon of diploidization, or is it just soaking up noise?
\end{quote}

%B: Rosana, you know this best, but it seems to me that the current phrasing (below) is unclear.
%R- rewritten
Adding hidden states to models more likely `soaks up' noise.
Therefore, in the presence of hidden states, finding that the diploidization parameter is different from zero truly represents potential evidence of the reversibility of polyploidy. 
Furthermore, all the Bayes factors indicated a moderate preference for models that include diploidization. %CHECK: I changed this because the rest of the manuscript states "moderate" and our tables show that the hidden state model shows "moderate" preference.
However, diploidization is a genomic process that may not be fully recovered from ploidy data. 
As discussed at the end of the article, genomic evidence in favor or against diploidization in Solanaceae is far from final.

@@

\begin{quote}
--I think this is an excellent point -- ``SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.''
Perhaps an discussion of the extent to which these results are extendable to other plant groups would be appropriate, versus what is likely to be specific to this system
\end{quote}

We include a necessarily limited discussion of other systems.

@@

\begin{quote}
--410  ``as self-incompatibility loci''?
\end{quote}

Fixed.

\begin{quote}
--417  Why would the rates for diploids be inflated?
So the diploids were ancestrally polyploidy -- why would that alter the rate inferences?
\end{quote}

We edited the discussion. 
We found no conclusive evidence about diversification changing in the presence of diploidization. 
The estimates for rate of polyploidization in the absence of diploidization also remained the similar (\cref{figure:hissedip}. 
However, ignoring diploidization, if it is common, could cause an underestimate of the polyploidization rate or create larger credible intervals.

We have corrected the phrase accordingly.
Thank you for the suggestion.
\begin{figure}
    \centering 
      \begin{minipage}[b]{0.5\textwidth}
    \includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf} 
    \end{minipage}\\
    \vspace{1cm}
    \begin{minipage}[b]{0.5\textwidth}
        \includegraphics[width=\textwidth]{hisseDPasymposteriordist.pdf}
        \end{minipage}
    \caption{Top six panels: ploidy model with hidden states without diploidization. Bottom six panels: ploidy model with hidden states with diploidization. Rate of polyploidy has a wider credible interval in the absence of diploidization.}
    \label{figure:hissedip}
\end{figure}

@@

We have now fitted models with asymmetric hidden rates and asymmetric trait rates. We found that the former are important and strengthened the differences. We found no evidence that having all asymmetric rates is preferable over just asymmetric hidden for our specific example (Figure S5, Table S6).

\begin{quote}
Finally, I have a suggestion for the authors.
One thing that I would have like to see is a discussion regarding the distribution of the hidden states at the tip of the tree (e.g., Fig S6, S8).
The two peaks distributions obtained for the hidden and non-hidden states made me wonder if the high and low rate peaks could be restricted to specific clades on the phylogeny.
After inspecting the supplementary material, I see that it is not the case; hidden states are rather spread throughout the phylogeny.
I think this is interesting and that it could make an interesting discussion.
\end{quote}

Thank you for this suggestion. 
After fitting the new models, the only trend we can notice is that for ploidy-only models, hidden states seem to transition mostly towards recent lineages but are still spread across the tree (Figure S8). 
For breeding system-only models, there are a two or three clades in which transitions between hidden states happen sooner (Figure S12 and S16). 
However, we do not have a clear interpretation for these phenomena.

### 2019-07-18 manuscript/discussion.tex
Oops! Fixing the opening quote and adding its reference! - adding final sentence emphsizing the same statement from honoree Barrett
Itay's comments related to discussion.tex & main.tex (#28-#32)

% E: should this results summary be much shorter?
Previous analyses of the effects of ploidy on diversification found that diploids are associated with greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}. 
We obtain a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification is not clear when we incorporate a hidden factor linked to diversification (\cref{figure:netdivall}B).
%
Previous analyses of breeding system in this family found that SI may cause higher diversification rates, compared with SC (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
Our analyses that include a hidden trait recover the same pattern, with one important difference.
On the background of one hidden state, we recover a net diversification rate for SC species that is positive and greater than the diversification rate of SI species on the background of the other hidden state (\cref{figure:netdivall}D).%B: I rewrote this to separate the point of the side clause. 
Therefore, SC may not be a `dead end' (cf.\ \citealt{igic_2013}) when coupled with some unknown trait combinations (or processes not modeled by these two traits). 
%Separately, I should note here that nothing about this analysis seems to me to strongly imply that SC is not a dead end in combination with a trait (as we discuss immediately below!). Much of the explanatory power derived for the A/B model may come from (for instance) accommodating tree scaling problems or rate heterogeneity in molecular evolution. It just fills up all of the mystery space without pointing to a cause. Side notes: (1) The I/C + A/B model ancestral states basically have the root of the tree as being in one hidden state and then tips in another. (2) Is there any place where we can see how much of the tree is in hidden state A or state B under some model?
%
Our analyses also reveal that models of joint evolution of ploidy and breeding system are statistically preferred, and hint at how various trait combinations may be linked with diversification.
We find that the highest net diversification rate is associated with SI diploids, while SC diploids have a lower diversification rate that overlaps with the net diversification of SC polyploids (\cref{figure:netdivall}E).
Thus, breeding system appears to provide a relatively better explanation of diversification rate differences, with ploidy providing a secondary effect within SC species.

@@

Throughout our numerous model comparisons, we find that inclusion of hidden states provides a considerably better fit (\cref{supptable:testaddhidden}). 
This is consistent with the expectation that many processes, beyond those associated with the focal traits, can affect inference of speciation and extinction. %removed "one or two" because it's true for N traits.
%A natural follow-up question is, what is this hidden trait?
It is, however, unclear exactly which processes are captured by the hidden states.
For example, our results show that, to a varying extent, breeding system functions as a hidden state in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}).
But the strong statistical support for processes not well explained by ploidy and breeding system (\cref{supptable:testaddhidden}) tempts one to interpret the remaining variation as the effect of other measurable traits.
For example, our data appear to show a rapidly-diversifying Australasian clade of mostly SC species within \emph{Solanum}, which suggests that geography may play a role. % Boris, please check!  Australian?  fixme
Nevertheless, it is also possible that the addition of hidden states instead explains variation stemming from any of a number of unrelated processes or methodological artifacts, as previously discussed by \citet{beaulieu_2016}.
In the absence of additional information, the hidden states can be viewed as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.

%such as variation in rates of molecular evolution or misspecified models of sequence evolution. %Rosana and Emma check me.
%R- let's not get into the weeds of what is a hidden state (I remove the use of trait  for state because a state is what the model receives) . Jeremy does a good job discussing this in his paper so no need.

%     but don't be circular, test in different clade
% alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model - it is low because not enough transitions and mostly a different process towards the tips-who knows what is happening there?
%B: this transition from the meaning of hidden states to best practices of lumping is a bit hard.- Fixed below
%Here Rosana adds the value of adding traits with the lumped models
As more trait information becomes available for macroevolutionary studies, it is important not only recognizing the role of hidden states as a part of a general modeling approach, but also questioning whether adding more traits to diversification studies is relevant.
More generally, when analyzing the value of adding additional traits to an already complex model with or without hidden states, we find that using lumped models  is a key tool to verify that adding a trait improves our understanding of the diversification patterns (\cref{figure:lumped,table:lumped}).

%R-removed this, it is really not important to discuss
%In our analyses, we find that the transition rates between and within hidden traits did not affect the direction of the net diversification differences. 
%We find that it is important to allow for the hidden state transition rates to be asymmetrical, because those models tend to provide a better fit than models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).

Although we fit an extensive set of models in order to relax a variety of assumptions, we did not explore the process of trait change in conjunction with speciation.
That is, our models all assume anagenetic trait evolution and ignore cladogenetic shifts.
Anagenetic and cladogenetic changes can be separated with phylogenetic models \citep{mayrose_2011, goldberg_2012, magnuson-ford_2012}. 
These have been previously applied to estimate the relative contribution of anagenetic and cladogenetic shifts in breeding system \citep{goldberg_2012} and polyploidization \citep{zhan_2016, freyman_2017}.
We did not explore cladogenetic trait change because of the vast increase in state space and model number, outside the scope of our already complex and extensive modeling framework. 
Although \citet{goldberg_2012} found that allowing cladogenetic changes did not substantially affect inference of net diversification rates associated with breeding system, future work could test whether this process affects diversification rate estimates with the more complex state spaces of our other models.

@@

Our findings prompt several further questions about the macroevolutionary pathways of ploidy and breeding system.
%
First, our support for the direct pathway is consistent with the idea that breakdown of SI by whole genome duplication---via diploid `heteroallelic' pollen---may often trigger the evolution of gender dimorphism as a different mechanism of inbreeding avoidance \citep{miller_2000}.
A further test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
%
Second, we might wonder whether the propensity for a polyploid species to diversify depends on whether it arose via the one-step or two-step pathway.
This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided, based on the path taken into that state.
This would be similar our analysis, but it would also include the possibility of different diversification rates in those $CP$ substates.
%
Third, the generality of our findings in other families remains to be assessed.
An identical procedure could be used in other families with gametophytic SI.
In clades with sporophytic SI systems, however, SI is not disabled by whole genome duplication, so there is no one-step pathway \citep{miller_2000,mable_2004}.
The correlation between breeding system and ploidy may therefore be different in sporophytic systems, and it is unclear whether one of the two-step pathways might predominate.

% pathways: compare/contrast with robertson2011
%     \rho_I > \rho_C in all!
%     likely way to lose SI
%     likely way to become P
%     (compare those relative rates with robertson2011)

@@

Polyploidization is known to be common in plants, but the relative frequency of the reverse process---diploidization---remains unclear, and it is under active investigation \citep{dodsworth_2015,mandakova_2018}.
%
Ignoring diploidization, if it is common, could cause underestimation and increase of uncertainty in  the polyploidization rate. 
Therefore, we included a slate of models with a diploidization parameter, and show that our main conclusions are robust to this process (\cref{suppfigure:alldip}).
These models also suggest modest statistical support for diploidization (\cref{supptable:testdiploidization}), although our estimates of its rate were highly uncertain. 
Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus-level chromosome multiplicity that we primarily relied on) are needed for more reliable conclusions.

@@

Heterogeneity in lineage diversification across time and clades is the rule, rather than the exception.
This background heterogeneity makes it difficult to test for the association of any one, isolated trait with different rates of speciation or extinction.
Our study provides an example of how diversification linked to a particular trait can be better assessed by a suite of more inclusive models that allow for alternative explanations---whether other traits or unknown factors.
Additionally, our analysis of evolutionary pathways to polyploidy shows the importance of including diversification effects even when addressing questions that focus on trait evolution.

>>>>

% E: should this results summary be much shorter?
Previous analyses of the effects of ploidy on diversification found that diploids are associated with greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}. 
We obtain a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification is not clear when we incorporate a hidden factor linked to diversification (\cref{figure:netdivall}B).
%
Previous analyses of breeding system in this family found that SI may cause higher diversification rates, compared with SC (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
Our analyses that include a hidden trait recover the same pattern, with one important difference.
On the background of one hidden state, we recover a net diversification rate for SC species that is positive and greater than the diversification rate of SI species on the background of the other hidden state (\cref{figure:netdivall}D). 
Therefore, SC may not be a `dead end' when coupled with some unknown trait combinations or processes not modeled by these two traits. 
%B: nothing about this analysis seems to me to strongly imply that SC is not a dead end in combination with a trait (as we discuss immediately below!). Much of the explanatory power derived for the A/B model may come from (for instance) accommodating tree scaling problems or rate heterogeneity in molecular evolution. It just fills up all of the mystery space without pointing to a cause. Side notes: (1) The I/C + A/B model ancestral states basically have the root of the tree as being in one hidden state and then tips in another. (2) Is there any place where we can see how much of the tree is in hidden state A or state B under some model?
%
Our analyses also reveal that models of joint evolution of ploidy and breeding system are statistically preferred, and hint at how various trait combinations may be linked with diversification.
We find that the highest net diversification rate is associated with SI diploids, while SC diploids have a lower diversification rate that overlaps with the net diversification of SC polyploids (\cref{figure:netdivall}E).
Thus, breeding system appears to provide a relatively better explanation of diversification rate differences, with ploidy providing a secondary effect within SC species.

@@

Throughout our numerous model comparisons, we find that inclusion of hidden states provides a considerably better fit (\cref{supptable:testaddhidden}). 
This is consistent with the expectation that many processes, beyond those associated with the focal traits, can affect inference of speciation and extinction. 
It is, however, unclear exactly which processes are captured by the hidden states. 
For example, our results show that, to a varying extent, breeding system functions as a hidden state in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}). 
But the strong statistical support for processes not well explained by ploidy and breeding system (\cref{supptable:testaddhidden}) tempts one to interpret the remaining variation as the effect of other measurable traits. 
For example, our data appear to show a rapidly-diversifying Australasian clade of mostly SC species within \emph{Solanum}, which suggests that geography may play a role. 
Nevertheless, it is also possible that the addition of hidden states instead explains variation stemming from any of a number of unrelated processes or methodological artifacts, as previously discussed by \citet{beaulieu_2016}.
In the absence of additional information, the hidden states can be viewed as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.

%R- let's not get into the weeds of what is a hidden state (I remove the use of trait  for state because a state is what the model receives). Jeremy does a good job discussing this in his paper so no need.
%     but don't be circular, test in different clade
% alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model - it is low because not enough transitions and mostly a different process towards the tips-who knows what is happening there?
%Here Rosana adds the value of adding traits with the lumped models
As more trait information becomes available for macroevolutionary studies, it is important not only recognizing the role of hidden states as a part of a general modeling approach, but also questioning whether adding more traits to diversification studies is relevant.
More generally, when analyzing the value of adding additional traits to an already complex model with or without hidden states, we find that using lumped models  is a key tool to verify that adding a trait improves our understanding of the diversification patterns (\cref{figure:lumped,table:lumped}).
%R-removed this: %In our analyses, we find that the transition rates between and within hidden traits did not affect the direction of the net diversification differences. 
%We find that it is important to allow for the hidden state transition rates to be asymmetrical, because those models tend to provide a better fit than models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).

Although we fit an extensive set of models in order to relax a variety of assumptions, we did not explore the process of trait change in conjunction with speciation.
That is, our models all assume anagenetic trait evolution and ignore cladogenetic shifts.
Anagenetic and cladogenetic changes can be separated with phylogenetic models \citep{mayrose_2011, goldberg_2012, magnuson-ford_2012}. 
These have been previously applied to estimate the relative contribution of anagenetic and cladogenetic shifts in breeding system \citep{goldberg_2012} and polyploidization \citep{zhan_2016, freyman_2017}.
We did not explore cladogenetic trait change because of the vast increase in parameter space and model number, outside the scope of our already complex and extensive modeling framework. %IM: #29
Although \citet{goldberg_2012} found that allowing cladogenetic changes did not substantially affect inference of net diversification rates associated with breeding system, future work could test whether this process affects diversification rate estimates with the more complex state and parameter spaces of our other models.

@@

Our findings prompt several further questions about the macroevolutionary pathways of ploidy and breeding system.
%
First, our support for the direct pathway is consistent with the idea that breakdown of SI by whole genome duplication---via diploid `heteroallelic' pollen---may often trigger the evolution of gender dimorphism as a different mechanism of inbreeding avoidance \citep{miller_2000}.
A further test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
Second, we might wonder whether the propensity for a polyploid species to diversify depends on whether it arose via the one-step or two-step pathway.
This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided into parts, reflecting the arriving path taken to that state. %IM: #30
Such an estimate of path-dependence could also include the possibility of different diversification rates for the subdivided $CP$ substates.
Third, the generality of our findings in other families remains to be assessed.
An identical procedure could be used in other families with gametophytic SI.
In clades with sporophytic SI systems, however, SI is not disabled by whole genome duplication, so there is no one-step pathway \citep{miller_2000,mable_2004}.
The correlation between breeding system and ploidy may therefore be different in sporophytic systems, and it is unclear whether one of the two-step pathways might predominate.
% pathways: compare/contrast with robertson2011
%     \rho_I > \rho_C in all!
%     likely way to lose SI
%     likely way to become P

@@

Polyploidization is known to be common in plants, but the pace and relative frequency of the reverse process---diploidization---remains unclear, and it is under active investigation \citep{dodsworth_2015,mandakova_2018}. %IM: #31
Ignoring diploidization, if it is common, could cause underestimation and increase of uncertainty in  the polyploidization rate. 
Therefore, we included a slate of models with a diploidization parameter, and show that our main conclusions are robust to this process (\cref{suppfigure:alldip}).
These models also suggest modest statistical support for diploidization (\cref{supptable:testdiploidization}), although our estimates of its rate were highly uncertain. 
Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus-level chromosome multiplicity that we primarily relied on) are needed for more reliable conclusions.

@@

Heterogeneity in lineage diversification across time and clades is the rule, rather than the exception.
This background heterogeneity makes it difficult to test for the association of any one, isolated trait with different rates of speciation or extinction.
Our study provides an example of how diversification linked to a particular trait can be better assessed by a suite of more inclusive models that allow for alternative explanations---whether other traits or unknown factors.
Additionally, our analysis of evolutionary pathways to polyploidy shows the importance of including diversification effects even when addressing questions that focus on trait evolution.
Finally, although a growing number of plant traits have been studied to date, breeding systems indeed seem to be among the most influential in governing macroevolution \citep{barrett1996}.

### 2019-07-18 manuscript/main.tex
Itay's comments related to discussion.tex & main.tex (#28-#32)

%--------------------------------------------------
\section{Acknowledgements}
%--------------------------------------------------
Special thanks to Bob Thomson for donating computational resources that made possible testing the convergence of the models. 
We also thank Sergei Tarasov for clarifying lumpability concepts, and Carrie Tribble for troubleshooting some ancestral estimate figures. 
The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
This work was supported by the National Science Foundation grant DEB-1655478 (to EEG), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Binational Science Foundation BSF-2016550 (to IM).
%B: new:
We thank Dan Schoen and three anonymous reviewers for providing uncommonly thoughtful and helpful reviews, as well as Spencer Barrett for continually inspiring our work.

>>>>

%--------------------------------------------------
\section{Acknowledgements}
%--------------------------------------------------
Special thanks to Bob Thomson for donating computational resources that made possible testing the convergence of the models. 
We also thank Sergei Tarasov for clarifying lumpability concepts, and Carrie Tribble for troubleshooting some ancestral estimate figures. 
The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
This work was supported by the National Science Foundation grant DEB-1655478 (to EEG), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Israel Science Foundation 961/17 (to IM). %IM: #32
%B: new:
We thank Dan Schoen and three anonymous reviewers for providing uncommonly thoughtful and helpful reviews, as well as Spencer Barrett for continually inspiring our work.
